---
title: "p8105_hw2_zz3167"
author: "Zhiyi Zhu"
date: "2023-09-29"
output: github_document
---

```{r,message=FALSE}
library(tidyverse)
```


## Problem 1

### Clean the data in pols-month.csv

```{r}
pols_df = 
  read_csv("data/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  mutate(
    month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
  ) |>
  mutate(
    president = case_match(
      prez_gop,
      1 ~ "gop",
      0 ~ "dem" )
  ) |>
  select(-prez_gop, -prez_dem, -day)
```


### Clean the data in snp.csv

```{r}
snp_df = 
  read_csv("data/snp.csv") |>
  janitor::clean_names() |>
  mutate(
    date = as.Date(date, format = "%m/%d/%y"),
    date = as.Date(ifelse(date > Sys.Date(),
                           format(date, "19%y-%m-%d"),
                           format(date)))
  ) |>
  separate(date, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  select(-day) |>
  arrange(year,month) |>
  mutate(
     month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
   ) |>
  select(year, month, everything())
```

### Clean the the data in unemployment.csv

```{r}
unem_df = 
  read_csv("data/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment percentage"
               )|>
  mutate(
    month = case_match(
      month,
      "jan" ~ "January",
      "feb" ~ "Febuary",
      "mar" ~ "March",
      "apr" ~ "April", 
      "may" ~ "May",
      "jun" ~ "June",
      "jul" ~ "July",
      "aug" ~ "August",
      "sep" ~ "September",
      "oct" ~ "October",
      "nov" ~ "November",
      "dec" ~ "December")
  )
```

### Join the datasets by merging snp into pols, and merging unemployment into the result

```{r}
snp_pols_df = 
  left_join(pols_df, snp_df, by = c("year", "month"))

merged_df = 
  left_join(snp_pols_df, unem_df, by = c("year", "month"))
```

### Write a short paragraph about these datasets

* The dataset `pols_df` has `r nrow(pols_df)` rows(observations) and `r ncol(pols_df)` columns(variables). The range of the year is from January 1947 to June 2015. The key variables are `r names(pols_df)`.
* The dataset `snp_df` has `r nrow(snp_df)` rows(observations) and `r ncol(snp_df)` columns(variables). The range of years is from January 1950 to July 2015. The key variables are `r names(snp_df)`.
* The dataset `unem_df` has `r nrow(unem_df)` rows(observations) and `r ncol(unem_df)` columns(variables). The range of years is from January 1948 to June 2015. The key variables are `r names(unem_df)`.
* The resulting dataset `merged_df` has `r nrow(merged_df)` rows(observations) and `r ncol(merged_df)` columns(variables). The range of years is from January 1947 to June 2015. The names of key variables are `r names(merged_df)`.













