---
title: "p8105_hw2_zz3167"
author: "Zhiyi Zhu"
date: "2023-09-29"
output: github_document
---

```{r,message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
```


## Problem 1

### Clean the data in pols-month.csv

```{r, message = FALSE}
pols_df = 
  read_csv("data/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  mutate(
    month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
  ) |>
  mutate(
    president = case_match(
      prez_gop,
      1 ~ "gop",
      0 ~ "dem" )
  ) |>
  select(-prez_gop, -prez_dem, -day)
```


### Clean the data in snp.csv

```{r, message = FALSE}
snp_df = 
  read_csv("data/snp.csv") |>
  janitor::clean_names() |>
  mutate(
    date = as.Date(date, format = "%m/%d/%y"),
    date = as.Date(ifelse(date > Sys.Date(),
                           format(date, "19%y-%m-%d"),
                           format(date)))
  ) |>
  separate(date, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  select(-day) |>
  arrange(year,month) |>
  mutate(
     month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
   ) |>
  select(year, month, everything())
```

### Clean the the data in unemployment.csv

```{r, message = FALSE}
unem_df = 
  read_csv("data/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment percentage"
               )|>
  mutate(
    month = case_match(
      month,
      "jan" ~ "January",
      "feb" ~ "Febuary",
      "mar" ~ "March",
      "apr" ~ "April", 
      "may" ~ "May",
      "jun" ~ "June",
      "jul" ~ "July",
      "aug" ~ "August",
      "sep" ~ "September",
      "oct" ~ "October",
      "nov" ~ "November",
      "dec" ~ "December")
  )
```

### Join the datasets by merging snp into pols, and merging unemployment into the result

```{r}
snp_pols_df = 
  left_join(pols_df, snp_df, by = c("year", "month"))

merged_df = 
  left_join(snp_pols_df, unem_df, by = c("year", "month"))
```

### Write a short paragraph about these datasets

* The dataset `pols_df` has `r nrow(pols_df)` rows(observations) and `r ncol(pols_df)` columns(variables). The range of the year is from January 1947 to June 2015. The key variables are `r names(pols_df)`.
* The dataset `snp_df` has `r nrow(snp_df)` rows(observations) and `r ncol(snp_df)` columns(variables). The range of years is from January 1950 to July 2015. The key variables are `r names(snp_df)`.
* The dataset `unem_df` has `r nrow(unem_df)` rows(observations) and `r ncol(unem_df)` columns(variables). The range of years is from January 1948 to June 2015. The key variables are `r names(unem_df)`.
* The resulting dataset `merged_df` has `r nrow(merged_df)` rows(observations) and `r ncol(merged_df)` columns(variables). The range of years is from January 1947 to June 2015. The names of key variables are `r names(merged_df)`.


## Problem 2

### Read and clean the Mr. Trash Wheel sheet

```{r}
MrTrash_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel", range = "A2:N586")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)
```

### Update the data to include a new homes_powered variable

```{r}
MrTrash_df = 
  mutate(
    MrTrash_df,
    homes_powered = weight_tons*500/30
      )
```

### Use a similar process to work with Professor Trash Wheel and Gwynnda

```{r}
ProfessorTrash_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet = "Professor Trash Wheel", range = "A2:M108")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)|>
  mutate(homes_powered = weight_tons*500/30)
```

```{r}
Gwynnda_df = 
  read_excel("data/202309 Trash Wheel Collection Data.xlsx",sheet = "Gwynnda Trash Wheel", range = "A2:L157")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)|>
  mutate(homes_powered = weight_tons*500/30)
```

### Add an additional variable to all datasets to keep track

```{r}
MrTrash_df = mutate(MrTrash_df, track = "Mr Trash wheel")
ProfessorTrash_df = mutate(ProfessorTrash_df, track = "Professor Trash Wheel")
Gwynnda_df = mutate(Gwynnda_df, track = "Gwynnda Trash wheel")
```

### Combine the two dataset with Mr. Trash Wheel and do some adjustment

```{r}
Combine_trash = 
  bind_rows(MrTrash_df, ProfessorTrash_df, Gwynnda_df)|>
  relocate(track, .before = year)|>
  relocate(homes_powered, .after = plastic_bags)
```


### Write a paragraph about these data

* The dataset `MrTrash_df` has `r nrow(MrTrash_df)` rows(observations) and `r ncol(MrTrash_df)` columns(variables). The key variables are `r names(MrTrash_df)`.
* The dataset `ProfessorTrash_df` has `r nrow(ProfessorTrash_df)` rows(observations) and `r ncol(ProfessorTrash_df)` columns(variables). The key variables are `r names(ProfessorTrash_df)`.
* The dataset `Gwynnda_df` has `r nrow(Gwynnda_df)` rows(observations) and `r ncol(Gwynnda_df)` columns(variables). The key variables are `r names(Gwynnda_df)`.
* The resulting dataset `Combine_trash` has `r nrow(Combine_trash)` rows(observations) and `r ncol(Combine_trash)` columns(variables). The key variables are `r names(Combine_trash)`.
* The total weight of trash collected by Professor Trash Wheel is `r sum(ProfessorTrash_df$weight_tons)` tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(subset(Gwynnda_df, month == "July" & year == "2021", cigarette_butts))`. 


## Problem 3

### Import, clean, and tidy the dataset of baseline demographics

```{r, message = FALSE}
baseline_df = 
  read_csv("data_mci/MCI_baseline.csv", skip = 1) |>
  janitor::clean_names()|>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      0 ~ "female")
  )|>
  mutate(
    apoe4 = case_match(
      apoe4,
      1 ~ "APOE4 carrier", 
      0 ~ "APOE4 non-carrier")
  )
  
standard_baseline_df = filter(baseline_df, age_at_onset > current_age | age_at_onset == '.')
```

### Discuss important steps in the import process

* We use read_csv function to import data, then we use skip = 1 function to skip the first row cause it's the given notes help us to understand what the numbers in each columns represent.
* We use mutate function to change the sex and apoe4 values according to the explanation in row 1.
* We use filter function to remove any participants who do not meet the stated inclusion criteria.

### Discuss relevant features of the dataset

* Of the initial `r nrow(baseline_df)` participants enrolled, `r nrow(standard_baseline_df)` met the required inclusion criteria, but `r nrow(baseline_df)-nrow(standard_baseline_df)` had MCI prior to recruitment and did not meet the recruitment criteria.
* Among participants who meet the criteria, `r nrow(subset(standard_baseline_df, age_at_onset != '.')) ` developed MCI.
* The average baseline age is `r mean(standard_baseline_df$current_age)`.
* `r round(nrow(subset(standard_baseline_df, sex == "female" & apoe4 == "APOE4 carrier"))/nrow(subset(standard_baseline_df, sex == "female"))*100, digits = 2)`% women in the study are APOE4 carriers.

###  Import, clean, and tidy the dataset of longitudinally observed biomarker values

```{r, message = FALSE}
amyloid_df = 
  read_csv("data_mci/mci_amyloid.csv", skip = 1) |>
  janitor::clean_names()|>
  rename(id = study_id)
```

### Comment on the steps

* We use read_csv function to import data, then we use skip = 1 function to skip the first row which is the explanation.
* We use rename function to change the "study_id" to "id" in order to easily combine or compare the two dataset.

### Discuss relevant features of the dataset

* There are `r nrow(amyloid_df)` rows(observations) and `r ncol(amyloid_df)` columns(variables) in the dataset `amyloid_df`. The names of variables are: `r names(amyloid_df)`.

### Check whether some participants appear in only the baseline or amyloid datasets

```{r}
only_baseline = anti_join(standard_baseline_df, amyloid_df, by = "id")
only_amyloid = anti_join(amyloid_df, standard_baseline_df, by = "id")
```


* There are `r nrow(only_baseline)` participants only appear in baseline dataset, their ids are `r only_baseline$id`.
* There are `r nrow(only_amyloid)` participants only appear in amyloid dataset, their ids are `r only_amyloid$id`.

### Combine the demographic and biomarker dataset

```{r}
combine_df = inner_join(standard_baseline_df, amyloid_df, by = "id")
```

### Briefly describe the resulting dataset

* There are `r nrow(combine_df)` participants appear in both dataset. `r nrow(subset(combine_df, sex == "male"))` of them are male, and `r round(nrow(subset(combine_df, sex == "male" & apoe4 == "APOE4 carrier"))/nrow(subset(combine_df, sex == "male"))*100, digits = 2)`% male  are APOE4 carriers. `r nrow(subset(combine_df, sex == "female"))` of them are female, and `r round(nrow(subset(combine_df, sex == "female" & apoe4 == "APOE4 carrier"))/nrow(subset(combine_df, sex == "female"))*100, digits = 2)`% female are APOE4 carriers.

### Export combined dataset as CSV 

```{r}
write_csv(combine_df, "data_mci/combine_dataset.csv")
```







