---
title: "p8105_hw2_zz3167"
author: "Zhiyi Zhu"
date: "2023-09-29"
output: github_document
---

```{r,message=FALSE}
library(tidyverse)
library(readxl)
```


## Problem 1

### Clean the data in pols-month.csv

```{r}
pols_df = 
  read_csv("data/pols-month.csv") |>
  janitor::clean_names() |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  mutate(
    month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
  ) |>
  mutate(
    president = case_match(
      prez_gop,
      1 ~ "gop",
      0 ~ "dem" )
  ) |>
  select(-prez_gop, -prez_dem, -day)
```


### Clean the data in snp.csv

```{r}
snp_df = 
  read_csv("data/snp.csv") |>
  janitor::clean_names() |>
  mutate(
    date = as.Date(date, format = "%m/%d/%y"),
    date = as.Date(ifelse(date > Sys.Date(),
                           format(date, "19%y-%m-%d"),
                           format(date)))
  ) |>
  separate(date, into = c("year", "month", "day"), sep = "-") |>
  mutate(
    year = as.numeric(year)
  ) |>
  select(-day) |>
  arrange(year,month) |>
  mutate(
     month = case_match(
      month,
      "01" ~ "January",
      "02" ~ "Febuary",
      "03" ~ "March",
      "04" ~ "Aprial",
      "05" ~ "May",
      "06" ~ "June",
      "07" ~ "July",
      "08" ~ "August",
      "09" ~ "September",
      "10" ~ "October",
      "11" ~ "November",
      "12" ~ "December" )
   ) |>
  select(year, month, everything())
```

### Clean the the data in unemployment.csv

```{r}
unem_df = 
  read_csv("data/unemployment.csv") |>
  janitor::clean_names() |>
  pivot_longer(jan:dec,
               names_to = "month",
               values_to = "unemployment percentage"
               )|>
  mutate(
    month = case_match(
      month,
      "jan" ~ "January",
      "feb" ~ "Febuary",
      "mar" ~ "March",
      "apr" ~ "April", 
      "may" ~ "May",
      "jun" ~ "June",
      "jul" ~ "July",
      "aug" ~ "August",
      "sep" ~ "September",
      "oct" ~ "October",
      "nov" ~ "November",
      "dec" ~ "December")
  )
```

### Join the datasets by merging snp into pols, and merging unemployment into the result

```{r}
snp_pols_df = 
  left_join(pols_df, snp_df, by = c("year", "month"))

merged_df = 
  left_join(snp_pols_df, unem_df, by = c("year", "month"))
```

### Write a short paragraph about these datasets

* The dataset `pols_df` has `r nrow(pols_df)` rows(observations) and `r ncol(pols_df)` columns(variables). The range of the year is from January 1947 to June 2015. The key variables are `r names(pols_df)`.
* The dataset `snp_df` has `r nrow(snp_df)` rows(observations) and `r ncol(snp_df)` columns(variables). The range of years is from January 1950 to July 2015. The key variables are `r names(snp_df)`.
* The dataset `unem_df` has `r nrow(unem_df)` rows(observations) and `r ncol(unem_df)` columns(variables). The range of years is from January 1948 to June 2015. The key variables are `r names(unem_df)`.
* The resulting dataset `merged_df` has `r nrow(merged_df)` rows(observations) and `r ncol(merged_df)` columns(variables). The range of years is from January 1947 to June 2015. The names of key variables are `r names(merged_df)`.


## Problem 2

### Read and clean the Mr. Trash Wheel sheet

```{r}
MrTrash_df = 
  read_excel("data/202207 Trash Wheel Collection Data.xlsx",sheet = "Mr. Trash Wheel", range = "A2:N549")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)
```

### Update the data to include a new homes_powered variable

```{r}
MrTrash_df = 
  mutate(
    MrTrash_df,
    homes_powered = weight_tons*500/30
      )
```

### Use a similar process to work with Professor Trash Wheel and Gwynnda

```{r}
ProfessorTrash_df = 
  read_excel("data/202207 Trash Wheel Collection Data.xlsx",sheet = "Professor Trash Wheel", range = "A2:M96")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)|>
  mutate(homes_powered = weight_tons*500/30)
```

```{r}
Gwynnda_df = 
  read_excel("data/202207 Trash Wheel Collection Data.xlsx",sheet = "Gwynnda Trash Wheel", range = "A2:K108")|>
  janitor::clean_names()|>
  separate(date, into = c("year", "month0", "day"), sep = "-")|>
  select(-month0)|>
  relocate(year, .before = month)|>
  mutate(homes_powered = weight_tons*500/30)
```

### Add an additional variable to all datasets to keep track

```{r}
MrTrash_df = mutate(MrTrash_df, track = "Mr Trash wheel")
ProfessorTrash_df = mutate(ProfessorTrash_df, track = "Professor Trash Wheel")
Gwynnda_df = mutate(Gwynnda_df, track = "Gwynnda Trash wheel")
```

### Combine the two dataset with Mr. Trash Wheel and do some adjustment

```{r}
Combine_trash = bind_rows(MrTrash_df, ProfessorTrash_df, Gwynnda_df)
Combine_trash = relocate(Combine_trash, track, .before = year)
Combine_trash = relocate(Combine_trash, homes_powered, .after = plastic_bags)
```

<mark> dumpster has more than one start

### Write a paragraph about these data

* The dataset `MrTrash_df` has `r nrow(MrTrash_df)` rows(observations) and `r ncol(MrTrash_df)` columns(variables). The key variables are `r names(MrTrash_df)`.
* The dataset `ProfessorTrash_df` has `r nrow(ProfessorTrash_df)` rows(observations) and `r ncol(ProfessorTrash_df)` columns(variables). The key variables are `r names(ProfessorTrash_df)`.
* The dataset `Gwynnda_df` has `r nrow(Gwynnda_df)` rows(observations) and `r ncol(Gwynnda_df)` columns(variables). The key variables are `r names(Gwynnda_df)`.
* The resulting dataset `Combine_trash` has `r nrow(Combine_trash)` rows(observations) and `r ncol(Combine_trash)` columns(variables). The key variables are `r names(Combine_trash)`.
* The total weight of trash collected by Professor Trash Wheel is `r sum(ProfessorTrash_df$weight_tons)` tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r sum(subset(Gwynnda_df, month == "July" & year == "2021")$cigarette_butts)`. 








